\section{Introduction}\label{sec:introduction}
The goal of this project is to implement a value analysis for a subset of the
eBPF language. This analysis aims to compute an approximation of the values of
registers and memory cells at each program point. The analysis is based on the
principles of abstract interpretation, specifically using an interval based
abstract domain. This report documents the design choices and implementation
details of the analysis. The project builds on the already implemented
eBPF-tools\cite{eBPF}. eBPF-tools is able to generate control-flow graphs
from the micro-eBPF language. These control-flow graphs are the foundation of
the implementation provided within this project.

\subsection{Abstract domain}
Throughout this project, I have used the infinite intervals to represent the
possible values of both registers and memory locations. That is to say, that an
abstract value is an interval $[l, u]$ where $l, u \in \{-\infty, \infty\}$.
These infinite intervals require the use of widening, followed by narrowing, to
produce usable results.

In the micro-eBPF language we have 11 registers (r0-r10) and a memory section
Mem, which contains 512 different cells. Throughout the implementation both
registers and memory cells were treated in the same manner. That is to say,
that a register/memory cell could take on any value in the interval mentioned
previously, or they could be some unreachable state denoted by $\bot$. Also,
the interval $\{-\infty, \infty\}$ itself is the representation of $\top$,
indicating that the register/memory cell can take on any value.

\subsection{Modelling program states}
The modelling of program states is best described by an example. Say we have
some arbitrary node, n, within the control flow graph. Now given some
transition $t$ from $n \to n'$ through some expression $exp$, then the state at
$n'$, namely $s'$, is given by:

\begin{align}
  s' &= (\llbracket exp \rrbracket s) \sqcup_{Itv} t' \label{eq:1}
\end{align}

Where $t'$ is any other transition, which also ends in n'. That is to say, that
the new state s' at node n' is given by the union of all the states that came
before and end in the current node. For this to work properly, we first
initialize all states to the neutral element of the interval union, which is
$\bot$.


\subsection{Abstract interpretation of expressions}
In general my implementation follows the abstract interpretation of expressions
seen in Miné\cite{mine}. However, Miné does not provide a definition for the
binary operators. In my case I was not able to get the time to implement the
binary operators. As such, if we ever see a binary operator, then we
immediately just return $\top$.

\subsubsection{Memory lookups}\label{sec:introductionMem}
For a memory lookup \texttt{Mem[ri]}, the abstract value of the register
\texttt{ri} determines the possible memory addresses to be read. Let $[l, u]$
denote the abstract interval of \texttt{ri}. The set of possible addresses is
$\{l, l+1, \dots, u\}$, meaning the result of the memory lookup is the union of
the abstract values at these addresses:
\begin{align}
  \text{Let } R(s)&= \bigsqcup_{Itv_{i \in [l, u]}} \llbracket \texttt{r[i]} \rrbracket s \\
  \llbracket \texttt{Mem[ri]} \rrbracket s &= \bigsqcup_{Itv_{j \in R(s)}} \texttt{Mem[j]}
\end{align}

This means, that there could be cases where we are writing to multiple memory
cells at the same time, if we are using a register, which at the current state
is not a constant value.


\subsection{Iteration strategy}
In this project the iteration strategy used is the work-set algorithm. This
method was used to avoid re-evaluating parts of the program, that were not
affected by the previous state. For more details on the specific implementation
see Section~\ref{sec:workset}.

The work-set algorithm was implemented with widening and narrowing. Widening
($\nabla$) is defined as in Mine\cite{mine}, i.e. we have that: 
\begin{align}\label{eq:widening}
  [l_1, u_1] \nabla [l_2, u_2] = [&\text{if } l_2 < l_1 \text{
  then } - \infty \text{ else } l_1, \\ 
  &\text{if } u_2 > u_1 \text{ then }
  \infty \text{ else } u_1] \nonumber
\end{align}

And Narrowing ($\Delta$) is defined as:
\begin{align}\label{eq:narrowing}
  [l_1, u_1] \Delta [l_2, u_2] = [&\text{if } l_1 = -\infty \text{
  then } l_2 \text{ else } l_1, \\ 
  &\text{if } u_1 = \infty \text{ then }
  u_2 \text{ else } u_1]\nonumber
\end{align}

To not overdo this, I included a counter for each node in the control flow
graph. If we reach the node x times, then we perform one widening step.
Afterwards, we perform some y amount of narrowing steps, but to avoid
infinitely looping, we limit y to some finite number.


